---
title: "Reanalysis: Hood et al. (2022)"
author: "Gidon Frischkorn"
format: html
embed-resources: TRUE
---

```{r SetUp}
#| message: false
#| warning: false
library(here)
library(dplyr)
library(AntiSaccade)
library(lme4)
library(lmerTest)
library(brms)
```


## Reproduction of Results from Hood et al. (2022)

Hood et al. (2022) investigated if the working memory capacity is more strongly related to performance in the anti- then in the pro-saccade task. For this, they first included performance in the automated OSPAN as a predictor of performance in the anti- and pro-saccade tasks. The below analysis do not exactly reproduce the results but closely math them (see below for details).

```{r Reproduction_Hood_Results}
Hood_2022_complete <- Hood_2022_behavioral %>% 
  filter(!is.na(ospan))

Hood_2022_complete$pc <- Hood_2022_complete$n_corr / Hood_2022_complete$n_trials

fit_lme <- lme4::lmer(pc ~ task*z_total + (1 | sub),
                      data = Hood_2022_complete)

anova(fit_lme)

Hood_2022_behavioral %>% 
  summarise(cor_ospan = cor(n_corr/n_trials, z_total, use = "complete.obs"),
            .by = task)
```

## Reanalysis using Bayesian GLM

From our perspective the more appropriate way to analyze this data is using a Generalized liner model assuming a binomial distribution of the data. We will use a Bayesian approach to estimate the model. Usually these models are estimated using a logistic link function to ensure that probability estimates are in the range between 0 and 1.

```{r Binomial_GLM}
#| message: false
#| warning: false
hood_formula <- bf(n_corr | trials(n_trials) ~ 1 + task*z_total + (1 + task | sub),
                   family = binomial)

# set appropriate constrasts for the task variable
Hood_2022_complete$task <- as.factor(Hood_2022_complete$task)
contrasts(Hood_2022_complete$task) <- bayestestR::contr.equalprior

# set priors
hood_priors <- prior(logistic(0,1), class = Intercept) +
  prior(logistic(0,0.5), class = b) 

# fit model including an intercept, main effects of task and ospan, and their interaction 
hood_fit <- brm(hood_formula, 
                data = Hood_2022_complete, 
                family = binomial,
                prior = hood_priors,
                sample_prior = TRUE,
                chains = 4, 
                cores = 4, 
                iter = 20000, 
                warmup = 1000, 
                control = list(adapt_delta = 0.99),
                file = here("output","fit_hood"),
                file_refit = "on_change",
                backend = "cmdstanr")

# print summary
summary(hood_fit)
```

An alternative specification of this model is to directly asses the performance in the pro- and anti-saccade task and seperately predict the performance by the OSPAN performance.

```{r Binomial_GLM_Alternative}
#| message: false
#| warning: false
# specify model formula for alternative model
hood_formula_alternative <- bf(n_corr | trials(n_trials) ~ 0 + task + task:z_total + (0 + task | sub),
                    family = binomial)

# set prior
hood_priors_alternative <- prior(logistic(0,1), class = b, coef = "taskanti") +
  prior(logistic(0,1), class = b, coef = "taskpro") +
  prior(logistic(0, 0.5), class = b, coef = "taskanti:z_total") + 
  prior(logistic(0, 0.5), class = b, coef = "taskpro:z_total")

# fit model
hood_fit_alternative <- brm(hood_formula_alternative, 
                            data = Hood_2022_complete, 
                            family = binomial,
                            prior = hood_priors_alternative,
                            sample_prior = TRUE,
                            chains = 4, 
                            cores = 4, 
                            iter = 20000, 
                            warmup = 1000, 
                            control = list(adapt_delta = 0.99),
                            file = here("output","fit_hood_alternative"),
                            file_refit = "on_change",
                            backend = "cmdstanr")

# print summary
summary(hood_fit_alternative)
```

We can use these results of both these models to calculate the strength of evidence for OSPAN as a predictor of performance in the anti- and pro-saccade task.

```{r HypothesisTests}
# calculate hypothesis tests
hyp_hood <- hypothesis(hood_fit, 
                       c(ospan_main = "z_total = 0",
                         ospan_int = "task1:z_total = 0"))
hyp_hood_alternative <- hypothesis(hood_fit_alternative, 
                                   c(ospan_anti = "taskanti:z_total = 0",
                                     ospan_pro = "taskpro:z_total = 0",
                                     ospan_diff = "taskanti:z_total = taskpro:z_total"))
```

For the first model, there is evidence for OSPAN prediciting performance irrespective of the task version, $BF_{10} = `r round(1/hyp_hood$hypothesis$Evid.Ratio[1],2)`$, however there is evidence against the interaction of OSPAN with the task predictor, $BF_{01} = `r round(hyp_hood$hypothesis$Evid.Ratio[2],2)`$. Thus, higher OSPAN performance leads to equal increases in performance for both the pro- and anti-saccade task.

```{r results_BGLM}
# print summary of hypothesis tests
hyp_hood
```

When looking at the evidence in the alternative specification of the BGLM. We find evidence for OSPAN predicting performance in the Anti-Saccade task, $BF_{10} = `r round( 1/hyp_hood_alternative$hypothesis$Evid.Ratio[1],2)`$, but inconclusive evidence with respect to OSPAN predicting performance in the Pro-Saccade task, $BF_{01} = `r round( hyp_hood_alternative$hypothesis$Evid.Ratio[2],2)`$. Nevertheless, when directly comparing how strongly OSPAN predicts performance in the pro- and anti-saccade task, there is strong evidence for these effects not differing, $BF_{01} = `r round(hyp_hood_alternative$hypothesis$Evid.Ratio[3],2)`$. This is a know issue, that the difference between credible and non-credible, is not necessarily credible itself. 

For this case, there are two reasons for there being no difference in the prediction of performance in the pro- and anti-saccade condition, but different strengths of evidence when separately evaluting if OSPAN predicts performance in the anti- or the pro-saccade tasks:

1. Individual differences in the pro-saccade task are measured less reliably due to ceiling effects.
2. On the logit scale there are actually larger individual differences in the pro- then the anti-saccade condition (see the SD estimates in the alternative specification of the Bayesian GLM)

The first point leads to wider posteriors for the estimated relationship between OSPAN and performance in the pro-saccade task compared to the anti-saccade condition. And both point one and two explain why the descriptive correlation between OSPAN and performance in the pro- and anti-saccade task differ.

```{r results_BGLM_alternative}
# print summary of hypothesis tests
hyp_hood_alternative
```
